---
title: "ISC_Tsimane_cytokine_github.Rmd"
output:
  word_document: default
  html_document:
    df_print: paged
---

Tsimane helminth and cytokine models and plots

```{r}
library(dplyr)
library(tidyverse)
library(plyr)
library(ggplot2)

# read in file
re_all = read.csv("https://github.com/indiaaugusta/Tsimane-Cytokines/blob/master/complete_file.csv", stringsAsFactors = FALSE) # stimulation data with serum cytokines, stimulation cytokines per medium, all parasite data, and eosinophil + all WBC data.

# eosinophil counts
mean(re_all$eos_st, na.rm = TRUE) # 2229.737
median(re_all$eos_st, na.rm = TRUE) # 1950

eosplot = re_all %>% ggplot(aes(eos_st)) + geom_histogram(binwidth = 200, fill = "darkseagreen", color = "black") + theme_classic() + geom_vline(aes(xintercept=mean(re_all$eos_st, na.rm = TRUE)), color="darkslateblue", linetype="dashed", size=1.8) + scale_x_continuous(breaks=seq(0,8000,1000)) + geom_vline(aes(xintercept=500), color="orange", linetype="dashed", size=1.8) + xlab("eosinophils")  + theme(text = element_text(size=20)) + labs(tag = "a")

# bin eosinophils
re_all$bin_eos_1500 = ifelse(re_all$eos_st > 1500, 1, 0)
re_all$bin_eos_3000 = ifelse(re_all$eos_st > 3000, 1, 0)

re_all_1500 = re_all[!is.na(re_all$bin_eos_1500),]
re_all_3000 = re_all[!is.na(re_all$bin_eos_3000),]

```

Stimulation data- LPS and H1N1 Cytokines as a function of eosinophils. Figures and models.

```{r}

# split by sex 

wom_all = re_all[re_all$male_stim == 0,]
men_all = re_all[re_all$male_stim == 1,]

######################################## LPS #############################################

######################### Women ######################### 

summary(glm(data = wom_all, log(IFNg_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.1b_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.2_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.4_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.5_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.6_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.7_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.8_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.10_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.12p70_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.13_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(GM.CSF_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(TNF.a_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))

######################### Men ######################### 

summary(glm(data = men_all, log(IFNg_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.1b_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.2_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.4_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.5_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.6_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.7_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.8_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.10_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.12p70_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.13_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(GM.CSF_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(TNF.a_stim_lps) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))

######################################## H1N1 #############################################

######################### Women ######################### 

summary(glm(data = wom_all, log(IFNg_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.1b_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.2_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.4_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.5_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.6_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.7_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.8_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.10_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.12p70_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(IL.13_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(GM.CSF_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(TNF.a_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))

######################### Men ######################### 

summary(glm(data = men_all, log(IFNg_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.1b_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.2_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.4_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.5_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.6_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.7_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.8_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.10_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.12p70_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(IL.13_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(GM.CSF_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(TNF.a_stim_h1n1) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))


```

Now we are making z-scores for each individual's Th1 and Th2 cytokine profiles, which are averages of the transformed z-scores of all of their cytokines within each suite. We do it for each condition separately.

Add 1 and then log, and then z-score

```{r}
# add 1 and then log before z-score
addlogfun = function(x){log(x+1)}
addlog = as.data.frame(apply(re_all[,2:66], 2, addlogfun))

# Z-score transformation
normFunc <- function(x){(x-mean(x, na.rm = T))/sd(x, na.rm = T)}
forz = apply(addlog[,1:65], 2, normFunc)
forz1 = as.data.frame(cbind(forz, re_all$vid, re_all$eos_st, re_all$male_stim))

# TH1 in LPS
lpsfirst_1 = forz1 %>% dplyr::select(V66, V67,V68, IFNg_stim_lps, IL.2_stim_lps)
lpsfirst_1 = lpsfirst_1 %>% dplyr::rename(vid = V66, eos_st = V67, male = V68)
lps_means_1 = lpsfirst_1 %>% dplyr::select(-c(vid, eos_st, male))
lps_means_1 = mutate_all(lps_means_1, function(x) as.numeric(as.character(x)))
lps_means_1$z_th1 = rowMeans(lps_means_1, na.rm = TRUE)
lps_z1 = cbind(lps_means_1, lpsfirst_1$vid, lpsfirst_1$eos_st, lpsfirst_1$male)
lps_z1 = lps_z1 %>% dplyr::rename(vid = 'lpsfirst_1$vid', eos_st = 'lpsfirst_1$eos_st', male = 'lpsfirst_1$male')
lps_z1$eos_st = as.numeric(as.character(lps_z1$eos_st))

# add metadata
new_meta = stim_nokids %>% dplyr::select(vid, hrich, prich, h_pres, p_pres, male_stim, stim_age, PSR, bmi,c_leucocitos, c_eosinophils, pregnant ) 

# merge
th1_lps = left_join(lps_z1, new_meta)
th1_lps = unique(th1_lps)

# split by sex
fem = th1_lps[th1_lps$male_stim == 0,]
maless = th1_lps[th1_lps$male_stim == 1,]

# model 
summary(glm(data = fem, z_th1 ~ bmi + stim_age + log(eos_st) + pregnant + log(c_leucocitos)))
summary(glm(data = maless, z_th1 ~ bmi + stim_age + log(eos_st) +  log(c_leucocitos)))

## now Th2 in LPS
lpsfirst_2 = forz1 %>% dplyr::select(V66, V67, IL.4_stim_lps,IL.5_stim_lps, IL.13_stim_lps)
lpsfirst_2 = lpsfirst_2 %>% dplyr::rename(vid = V66, eos_st = V67)
row_means_2 = lpsfirst_2 %>% dplyr::select(-c(vid, eos_st))
row_means_2 = mutate_all(row_means_2, function(x) as.numeric(as.character(x)))
row_means_2$z_th2 = rowMeans(row_means_2)
lps_z_th2 = cbind(row_means_2, lpsfirst_2$vid, lpsfirst_2$eos_st)
lps_z_th2 = lps_z_th2 %>% dplyr::rename(vid = 'lpsfirst_2$vid', eos_st = 'lpsfirst_2$eos_st')
lps_z_th2$eos_st = as.numeric(as.character(lps_z_th2$eos_st))

# add metadata
new_meta = stim_nokids %>% dplyr::select(vid, hrich, prich, h_pres, p_pres, male_stim, stim_age, PSR, bmi,c_leucocitos, c_eosinophils, pregnant ) 

# merge
th2_lps = left_join(lps_z_th2, new_meta)
th2_lps = unique(th2_lps)

# split by sex
fem2 = th2_lps[th2_lps$male_stim == 0,]
male2 = th2_lps[th2_lps$male_stim == 1,]

# model 
summary(glm(data = fem2, z_th2 ~ bmi + stim_age + log(eos_st) + pregnant + log(c_leucocitos)))
summary(glm(data = male2, z_th2 ~ bmi + stim_age + log(eos_st) + log(c_leucocitos)))

# Th1 in H1N1
hn_1 = forz1 %>% dplyr::select(V66, V67,IFNg_stim_h1n1, IL.2_stim_h1n1)
hn_1 = hn_1 %>% dplyr::rename(vid = V66, eos_st = V67)
h1n1_means_1 = hn_1 %>% dplyr::select(-c(vid, eos_st))
h1n1_means_1 = mutate_all(h1n1_means_1, function(x) as.numeric(as.character(x)))
h1n1_means_1$z_th1 = rowMeans(h1n1_means_1)
h1n1_z_th1 = cbind(h1n1_means_1, hn_1$vid, hn_1$eos_st)
h1n1_z_th1 = h1n1_z_th1 %>% dplyr::rename(vid = 'hn_1$vid', eos_st = 'hn_1$eos_st')
h1n1_z_th1$eos_st = as.numeric(as.character(h1n1_z_th1$eos_st))

# add metadata
new_meta = stim_nokids %>% dplyr::select(vid, hrich, prich, h_pres, p_pres, male_stim, stim_age, PSR, bmi,c_leucocitos, c_eosinophils, pregnant ) 

# merge
th1_h1n1 = left_join(h1n1_z_th1, new_meta)
th1_h1n1 = unique(th1_h1n1)

# split by sex
fem_h1n1_1 = th1_h1n1[th1_h1n1$male_stim == 0,]
man_h1n1_1 = th1_h1n1[th1_h1n1$male_stim == 1,]

# model 
summary(glm(data = fem_h1n1_1, z_th1 ~ bmi + stim_age + log(eos_st) + pregnant + log(c_leucocitos)))
summary(glm(data = man_h1n1_1, z_th1 ~ bmi + stim_age + log(eos_st) + log(c_leucocitos)))

## now Th2 in H1n1
h1n1_2 = forz1 %>% dplyr::select(V66, V67, IL.4_stim_h1n1,IL.5_stim_h1n1, IL.13_stim_h1n1)
h1n1_2 = h1n1_2 %>% dplyr::rename(vid = V66, eos_st = V67)
h1n1_means_2 = h1n1_2 %>% dplyr::select(-c(vid, eos_st))
h1n1_means_2 = mutate_all(h1n1_means_2, function(x) as.numeric(as.character(x)))
h1n1_means_2$z_th2 = rowMeans(h1n1_means_2)
h1n1_z_th2 = cbind(h1n1_means_2, h1n1_2$vid, h1n1_2$eos_st)
h1n1_z_th2 = h1n1_z_th2 %>% dplyr::rename(vid = 'h1n1_2$vid', eos_st = 'h1n1_2$eos_st')
h1n1_z_th2$eos_st = as.numeric(as.character(h1n1_z_th2$eos_st))

# add metadata
new_meta = stim_nokids %>% dplyr::select(vid, hrich, prich, h_pres, p_pres, male_stim, stim_age, PSR, bmi,c_leucocitos, c_eosinophils, pregnant ) 

# merge
th2_h1n1 = left_join(h1n1_z_th2, new_meta)
th2_h1n1 = unique(th2_h1n1)

# split by sex
fem_h1n1_2 = th2_h1n1[th2_h1n1$male_stim == 0,]
man_h1n1_2 = th2_h1n1[th2_h1n1$male_stim == 1,]

# model 
summary(glm(data = fem_h1n1_2, z_th2 ~ bmi + stim_age + log(eos_st) + pregnant + log(c_leucocitos)))
summary(glm(data = man_h1n1_2, z_th2 ~ bmi + stim_age + log(eos_st) + log(c_leucocitos)))

```

Now we make forest plots for all the iterations

```{r}
# create forest plot df. cytokine, beta, beta+1.96*se, beta-1.96*se
allforest = read.csv("~/Desktop/Tsimane/all_forest.csv", stringsAsFactors = FALSE)
allforest_redo = read.csv("~/Desktop/Tsimane/all_forest_redo.csv", stringsAsFactors = FALSE)

order = c( 'LPS', 'H1N1')
allforest_redo$Type = factor(allforest_redo$Type, levels=order)

order1 = c( 'Male', 'Female')
allforest_redo$Sex = factor(allforest_redo$Sex, levels=order1)

dotCOLS = c("#f9b282", "#a6d8f0")
barCOLS = c("#de6b35", "#008fd5")

ggplot(allforest_redo, aes(x=Type, y=Estimate, ymin=Lower_CI, ymax=Upper_CI, color = Sex, fill = Sex))+
 geom_linerange(size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS, guide=TRUE)+
scale_color_manual(values=dotCOLS, guide=FALSE)+
scale_x_discrete(name="") +
scale_y_continuous(name="Eosinophil Effect (95% CI)", limits = c(-3, 3)) +
coord_flip() +
theme_classic(base_size=18 ) +
  facet_wrap(~Set, ncol = 1, ) + theme(strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=18), legend.title = element_blank()) + guides(fill = guide_legend(override.aes = list(linetype=0), reverse = TRUE))

```

Now we do heatmap

```{r}
heatall = read.csv("~/Desktop/Tsimane/heatmap_lps.csv", stringsAsFactors = FALSE)

# fix cytokine names
heatall$Cytokine = ifelse(heatall$Cytokine == "IFNg", "IFN\u03B3", heatall$Cytokine)
heatall$Cytokine = ifelse(heatall$Cytokine == "1b", "IL-1\u03B2", heatall$Cytokine)
heatall$Cytokine = ifelse(heatall$Cytokine == "TNF", "TNF-\u03B1", heatall$Cytokine)
heatall$Cytokine = ifelse(heatall$Cytokine == "GM", "GM-CSF", heatall$Cytokine)

# define colors and order
dotCOLS_1 = c( "#f9b282","#a6d8f0")
barCOLS_1 = c("#de6b35", "#008fd5" )

order5 = c( 'TNF-α', 'GM-CSF','IL-12p70', 'IL-10', 'IL-8', 'IL-7', 'IL-6', 'IL-1β','IL-13','IL-5', 'IL-4', 'IL-2', 'IFNγ')
heatall$Cytokine = factor(heatall$Cytokine, levels=order5)

order4 = c(  'Male', 'Female')
heatall$Sex = factor(heatall$Sex, levels=order4)

# make categories th1/th2/ambiguous
heatall$type = ifelse(heatall$Cytokine == "IFNγ" | heatall$Cytokine == "IL-2", "Th1", NA)
heatall$type = ifelse(heatall$Cytokine == "IL-4" | heatall$Cytokine == "IL-5" | heatall$Cytokine == "IL-13", "Th2", "Ambiv")

# plot

ggplot() +
  geom_rect(data=ToothGrowth,xmin = 0, xmax = 8.5, ymin = -7, ymax = 3, fill = '#e6e8e8', alpha = 0.01) + geom_rect(data=ToothGrowth,xmin = 8.5, xmax = 11.5, ymin = -7, ymax = 3, fill = '#fae8c8', alpha = 0.01) + geom_rect(data=ToothGrowth,xmin = 11.5, xmax = 14, ymin = -7, ymax = 3, fill = '#eddff5', alpha = 0.05) +
#geom_rect(data=ToothGrowth,xmin = 0, xmax = 8.5, ymin = -7, ymax = 2, fill = '#cbd5e8', alpha = 0.08) + geom_rect(data=ToothGrowth,xmin = 8.5, xmax = 11.5, ymin = -7, ymax = 2, fill = '#fdcdac', alpha = 0.08) + geom_rect(data=ToothGrowth,xmin = 11.5, xmax = 14, ymin = -7, ymax = 2, fill = '#b3e2cd', alpha = 0.08) +
geom_text(data=data.frame(0),x=08.25,y=-6.5,hjust=0,label='Other') +
geom_text(data=data.frame(0),x=11.25,y=-6.5,hjust=0,label='Th2') +
geom_text(data=data.frame(0),x=13.25,y=-6.5,hjust=0,label='Th1') +
geom_linerange(data=heatall, aes(x=Cytokine, ymin=Lower_CI, ymax=Upper_CI, color = Sex),size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(data=heatall, aes(x=Cytokine, y=Estimate, fill = Sex),size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS_1, guide=TRUE)+
scale_color_manual(values=dotCOLS_1, guide=FALSE)+
scale_y_continuous(name="Eosinophil Effect (95% CI)", limits = c(-7, 2), breaks=seq(-6,2,2)) +
theme_classic(base_size=18 ) +
facet_wrap(~Medum, ncol = 2, ) + theme(axis.title.y = element_blank(), strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=20), legend.title = element_blank(),axis.text.x = element_text(vjust =1, hjust=0, size = 16)) + guides(fill = guide_legend(override.aes = list(linetype=0), reverse = TRUE)) + coord_flip() 
```

Make forest plots for additional variables.

```{r}
allvar_forest = read.csv("~/Desktop/Tsimane/All_variables_forest.csv", stringsAsFactors = FALSE)

# special characterss
allvar_forest$Cytokine = ifelse(allvar_forest$Cytokine == "IFNg", "IFN\u03B3", allvar_forest$Cytokine)
allvar_forest$Cytokine = ifelse(allvar_forest$Cytokine == "1b", "IL-1\u03B2", allvar_forest$Cytokine)
allvar_forest$Cytokine = ifelse(allvar_forest$Cytokine == "TNF", "TNF-\u03B1", allvar_forest$Cytokine)
allvar_forest$Cytokine = ifelse(allvar_forest$Cytokine == "GM", "GM-CSF", allvar_forest$Cytokine)

# order variables
order3 = c( 'TNF-α', 'GM-CSF', 'IL-13', 'IL-12p70', 'IL-10', 'IL-8', 'IL-7', 'IL-6', 'IL-5', 'IL-4', 'IL-2','IL-1β', 'IFNγ')
allvar_forest$Cytokine = factor(allvar_forest$Cytokine, levels=order3)

order4 = c(  'M', 'F')
allvar_forest$Sex = factor(allvar_forest$Sex, levels=order4)

# BMI 
bmi_forest = allvar_forest[allvar_forest$Predictor == "BMI",]

dotCOLS_bmi = c("#f9b282", "#a6d8f0")
barCOLS_bmi = c("#de6b35", "#008fd5")

bmi = ggplot() +
geom_rect(data=ToothGrowth,xmin = 0, xmax = 8.5, ymin = -7, ymax = 3, fill = '#e6e8e8', alpha = 0.03) + geom_rect(data=ToothGrowth,xmin = 8.5, xmax = 11.5, ymin = -7, ymax = 3, fill = '#fdcdac', alpha = 0.01) + geom_rect(data=ToothGrowth,xmin = 11.5, xmax = 14, ymin = -7, ymax = 3, fill = '#b3e2cd', alpha = 0.01) +
geom_text(data=data.frame(0),x=08.25,y=-6.5,hjust=0,label='Other') +
geom_text(data=data.frame(0),x=11.25,y=-6.5,hjust=0,label='Th2') +
geom_text(data=data.frame(0),x=13.25,y=-6.5,hjust=0,label='Th1') +
geom_linerange(data=bmi_forest, aes(x=Cytokine, ymin=Lower_CI, ymax=Upper_CI, color = Sex),size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(data=bmi_forest, aes(x=Cytokine, y=Estimate, fill = Sex),size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS_bmi, guide=TRUE)+
scale_color_manual(values=dotCOLS_bmi, guide=FALSE)+
scale_y_continuous(name="BMI Effect (95% CI)", limits = c(-7, 2), breaks=seq(-6,2,2)) + 
theme_classic(base_size=18 ) +
facet_wrap(~Medum, ncol = 2, ) + theme(axis.title.y = element_blank(), strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=20), legend.title = element_blank(),axis.text.x = element_text(vjust =1, hjust=0, size = 16)) + guides(fill = guide_legend(override.aes = list(linetype=0), reverse = TRUE)) + coord_flip() + labs(tag = "a")

# Age 
age_forest = allvar_forest[allvar_forest$Predictor == "Age (continuous)",]

age = ggplot() +
geom_rect(data=ToothGrowth,xmin = 0, xmax = 8.5, ymin = -7, ymax = 3, fill = '#e6e8e8', alpha = 0.03) + geom_rect(data=ToothGrowth,xmin = 8.5, xmax = 11.5, ymin = -7, ymax = 3, fill = '#fdcdac', alpha = 0.01) + geom_rect(data=ToothGrowth,xmin = 11.5, xmax = 14, ymin = -7, ymax = 3, fill = '#b3e2cd', alpha = 0.01) +
geom_text(data=data.frame(0),x=08.25,y=-6.5,hjust=0,label='Other') +
geom_text(data=data.frame(0),x=11.25,y=-6.5,hjust=0,label='Th2') +
geom_text(data=data.frame(0),x=13.25,y=-6.5,hjust=0,label='Th1') +
geom_linerange(data=age_forest, aes(x=Cytokine, ymin=Lower_CI, ymax=Upper_CI, color = Sex),size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(data=age_forest, aes(x=Cytokine, y=Estimate, fill = Sex),size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS_bmi, guide=FALSE)+
scale_color_manual(values=dotCOLS_bmi, guide=FALSE)+
scale_y_continuous(name="Age Effect (95% CI)", limits = c(-7, 2), breaks=seq(-6,2,2)) + 
theme_classic(base_size=18 ) +
facet_wrap(~Medum, ncol = 2, ) + theme(axis.title.y = element_blank(), strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=20), legend.title = element_blank(),axis.text.x = element_text(vjust =1, hjust=0, size = 16)) + coord_flip()  + labs(tag = "b")

# Pregnancy 
preg_forest = allvar_forest[allvar_forest$Predictor == "Pregnant (Y)",]
dotCOLS_preg = c("#a6d8f0", "#f9b282")
barCOLS_preg = c("#008fd5", "#de6b35")

preg = ggplot() +
geom_rect(data=ToothGrowth,xmin = 0, xmax = 8.5, ymin = -7, ymax = 3, fill = '#e6e8e8', alpha = 0.03) + geom_rect(data=ToothGrowth,xmin = 8.5, xmax = 11.5, ymin = -7, ymax = 3, fill = '#fdcdac', alpha = 0.01) + geom_rect(data=ToothGrowth,xmin = 11.5, xmax = 14, ymin = -7, ymax = 3, fill = '#b3e2cd', alpha = 0.01) +
geom_text(data=data.frame(0),x=08.25,y=-6.5,hjust=0,label='Other') +
geom_text(data=data.frame(0),x=11.25,y=-6.5,hjust=0,label='Th2') +
geom_text(data=data.frame(0),x=13.25,y=-6.5,hjust=0,label='Th1') +
geom_linerange(data=preg_forest, aes(x=Cytokine, ymin=Lower_CI, ymax=Upper_CI, color = Sex),size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(data=preg_forest, aes(x=Cytokine, y=Estimate, fill = Sex),size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS_preg, guide=FALSE)+
scale_color_manual(values=dotCOLS_preg, guide=FALSE)+
scale_y_continuous(name="Pregnancy Effect (95% CI)", limits = c(-7, 3), breaks=seq(-6,2,2)) + 
theme_classic(base_size=18 ) +
facet_wrap(~Medum, ncol = 2, ) + theme(axis.title.y = element_blank(), strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=20), legend.title = element_blank(),axis.text.x = element_text(vjust =1, hjust=0, size = 16)) + coord_flip()  + labs(tag = "d")

# Leukocytes 
leuk_forest = allvar_forest[allvar_forest$Predictor == "Leukocyte count (logged)",]

leuk = ggplot() +
geom_rect(data=ToothGrowth,xmin = 0, xmax = 8.5, ymin = -7, ymax = 3, fill = '#e6e8e8', alpha = 0.03) + geom_rect(data=ToothGrowth,xmin = 8.5, xmax = 11.5, ymin = -7, ymax = 3, fill = '#fdcdac', alpha = 0.01) + geom_rect(data=ToothGrowth,xmin = 11.5, xmax = 14, ymin = -7, ymax = 3, fill = '#b3e2cd', alpha = 0.01) +
geom_text(data=data.frame(0),x=08.25,y=-6.5,hjust=0,label='Other') +
geom_text(data=data.frame(0),x=11.25,y=-6.5,hjust=0,label='Th2') +
geom_text(data=data.frame(0),x=13.25,y=-6.5,hjust=0,label='Th1') +
geom_linerange(data=leuk_forest, aes(x=Cytokine, ymin=Lower_CI, ymax=Upper_CI, color = Sex),size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(data=leuk_forest, aes(x=Cytokine, y=Estimate, fill = Sex),size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS_bmi, guide=FALSE)+
scale_color_manual(values=dotCOLS_bmi, guide=FALSE)+
scale_y_continuous(name="Leukocyte Effect (95% CI)", limits = c(-7, 3), breaks=seq(-6,2,2)) + 
theme_classic(base_size=18 ) +
facet_wrap(~Medum, ncol = 2, ) + theme(axis.title.y = element_blank(), strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=20), legend.title = element_blank(),axis.text.x = element_text(vjust =1, hjust=0, size = 16))  + guides(fill = guide_legend(override.aes = list(linetype=0), reverse = TRUE)) + coord_flip()  + labs(tag = "c")

# put all these bbs into single figure
library(ggpubr)

pdf(file="~/Desktop/Tsimane/othervariable_onepage.pdf",useDingbats=TRUE,height=20,width=20) 
ggarrange(bmi, age, leuk, preg, ncol=2,nrow=2,widths=c(2,2, 2,2, 2,2, 2,2),newpage=FALSE)
dev.off()
```

Define pro- and anti-inflammatory categories

```{r}
# add 1 and then log before z-score
addlogfun = function(x){log(x+1)}
addlog = as.data.frame(apply(re_all[,2:66], 2, addlogfun))

# Z-score transformation
normFunc <- function(x){(x-mean(x, na.rm = T))/sd(x, na.rm = T)}
forz = apply(addlog[,1:65], 2, normFunc)
forz1 = as.data.frame(cbind(forz, re_all$vid, re_all$eos_st, re_all$male_stim))

# pro in LPS
lpsfirst_pro = forz1 %>% dplyr::select(V66, V67,V68, IFNg_stim_lps, IL.2_stim_lps, IL.6_stim_lps, IL.8_stim_lps, IL.12p70_stim_lps, TNF.a_stim_lps)
lpsfirst_pro = lpsfirst_pro %>% dplyr::rename(vid = V66, eos_st = V67, male = V68)
lps_means_pro = lpsfirst_pro %>% dplyr::select(-c(vid, eos_st, male))
lps_means_pro = mutate_all(lps_means_pro, function(x) as.numeric(as.character(x)))
lps_means_pro$z_th1 = rowMeans(lps_means_pro, na.rm = TRUE)
lps_z_pro = cbind(lps_means_pro, lpsfirst_pro$vid, lpsfirst_pro$eos_st, lpsfirst_pro$male)
lps_z_pro = lps_z_pro %>% dplyr::rename(vid = 'lpsfirst_pro$vid', eos_st = 'lpsfirst_pro$eos_st', male = 'lpsfirst_pro$male')
lps_z_pro$eos_st = as.numeric(as.character(lps_z_pro$eos_st))

# add metadata
new_meta = stim_nokids %>% dplyr::select(vid, hrich, prich, h_pres, p_pres, male_stim, stim_age, PSR, bmi,c_leucocitos, c_eosinophils, pregnant ) 

# merge
lps_pro = left_join(lps_z_pro, new_meta)
lps_pro = unique(lps_pro)

# split by sex
fem_pro = lps_pro[lps_pro$male_stim == 0,]
maless_pro = lps_pro[lps_pro$male_stim == 1,]

# model 
summary(glm(data = fem_pro, z_th1 ~ bmi + stim_age + log(eos_st) + pregnant + log(c_leucocitos)))
summary(glm(data = maless_pro, z_th1 ~ bmi + stim_age + log(eos_st) +  log(c_leucocitos)))

## now anti in LPS
lpsfirst_anti = forz1 %>% dplyr::select(V66, V67, IL.4_stim_lps,IL.5_stim_lps, IL.10_stim_lps, IL.13_stim_lps)
lpsfirst_anti = lpsfirst_anti %>% dplyr::rename(vid = V66, eos_st = V67)
row_means_anti = lpsfirst_anti %>% dplyr::select(-c(vid, eos_st))
row_means_anti = mutate_all(row_means_anti, function(x) as.numeric(as.character(x)))
row_means_anti$z_th2 = rowMeans(row_means_anti)
lps_z_anti = cbind(row_means_anti, lpsfirst_anti$vid, lpsfirst_anti$eos_st)
lps_z_anti = lps_z_anti %>% dplyr::rename(vid = 'lpsfirst_anti$vid', eos_st = 'lpsfirst_anti$eos_st')
lps_z_anti$eos_st = as.numeric(as.character(lps_z_anti$eos_st))

# add metadata
new_meta = stim_nokids %>% dplyr::select(vid, hrich, prich, h_pres, p_pres, male_stim, stim_age, PSR, bmi,c_leucocitos, c_eosinophils, pregnant ) 

# merge
anti_lps = left_join(lps_z_anti, new_meta)
anti_lps = unique(anti_lps)

# split by sex
fem_anti = anti_lps[anti_lps$male_stim == 0,]
male_anti = anti_lps[anti_lps$male_stim == 1,]

# model
summary(glm(data = fem_anti, z_th2 ~ bmi + stim_age + log(eos_st) + pregnant + log(c_leucocitos)))
summary(glm(data = male_anti, z_th2 ~ bmi + stim_age + log(eos_st) + log(c_leucocitos)))

# pro in H1N1
hn_pro = forz1 %>% dplyr::select(V66, V67,IFNg_stim_h1n1, IL.2_stim_h1n1)
hn_pro = hn_pro %>% dplyr::rename(vid = V66, eos_st = V67)
h1n1_means_pro = hn_pro %>% dplyr::select(-c(vid, eos_st))
h1n1_means_pro = mutate_all(h1n1_means_pro, function(x) as.numeric(as.character(x)))
h1n1_means_pro$z_th1 = rowMeans(h1n1_means_pro)
h1n1_z_pro = cbind(h1n1_means_pro, hn_pro$vid, hn_pro$eos_st)
h1n1_z_pro = h1n1_z_pro %>% dplyr::rename(vid = 'hn_pro$vid', eos_st = 'hn_pro$eos_st')
h1n1_z_pro$eos_st = as.numeric(as.character(h1n1_z_pro$eos_st))

# add metadata
new_meta = stim_nokids %>% dplyr::select(vid, hrich, prich, h_pres, p_pres, male_stim, stim_age, PSR, bmi,c_leucocitos, c_eosinophils, pregnant ) 

# merge
pro_h1n1 = left_join(h1n1_z_pro, new_meta)
pro_h1n1 = unique(pro_h1n1)

# split by sex
man_h1n1_pro = pro_h1n1[pro_h1n1$male_stim == 1,]
fem_h1n1_pro = pro_h1n1[pro_h1n1$male_stim == 0,]

# model
summary(glm(data = fem_h1n1_pro, z_th1 ~ bmi + stim_age + log(eos_st) + pregnant + log(c_leucocitos))) 
summary(glm(data = man_h1n1_pro, z_th1 ~ bmi + stim_age + log(eos_st) + log(c_leucocitos)))

## now anti in H1n1
h1n1_anti = forz1 %>% dplyr::select(V66, V67, IL.4_stim_h1n1,IL.5_stim_h1n1,IL.10_stim_h1n1, IL.13_stim_h1n1)
h1n1_anti = h1n1_anti %>% dplyr::rename(vid = V66, eos_st = V67)
h1n1_means_anti = h1n1_anti %>% dplyr::select(-c(vid, eos_st))
h1n1_means_anti = mutate_all(h1n1_means_2, function(x) as.numeric(as.character(x)))
h1n1_means_anti$z_th2 = rowMeans(h1n1_means_anti)
h1n1_z_anti = cbind(h1n1_means_anti, h1n1_anti$vid, h1n1_anti$eos_st)
h1n1_z_anti = h1n1_z_anti %>% dplyr::rename(vid = 'h1n1_anti$vid', eos_st = 'h1n1_anti$eos_st')
h1n1_z_anti$eos_st = as.numeric(as.character(h1n1_z_anti$eos_st))

# add metadata
new_meta = stim_nokids %>% dplyr::select(vid, hrich, prich, h_pres, p_pres, male_stim, stim_age, PSR, bmi,c_leucocitos, c_eosinophils, pregnant ) 

# merge
h1n1_z_anti = left_join(h1n1_z_anti, new_meta)
h1n1_z_anti = unique(h1n1_z_anti)

# split by sex
fem_h1n1_anti = h1n1_z_anti[h1n1_z_anti$male_stim == 0,]
man_h1n1_anti = h1n1_z_anti[h1n1_z_anti$male_stim == 1,]

# model
summary(glm(data = fem_h1n1_anti, z_th2 ~ bmi + stim_age + log(eos_st) + pregnant + log(c_leucocitos)))
summary(glm(data = man_h1n1_anti, z_th2 ~ bmi + stim_age + log(eos_st) + log(c_leucocitos)))

# now make tables for these and then make supplementary forest plots for these ones

# create forest plot df. cytokine, beta, beta+1.96*se, beta-1.96*se
suppforest = read.csv("~/Desktop/Tsimane/Supp_forest.csv", stringsAsFactors = FALSE)
suppforest = suppforest[!(is.na(suppforest$Type)),]
suppforest_redo = read.csv("~/Desktop/Tsimane/Supp_forest_redo.csv", stringsAsFactors = FALSE)
suppforest_redo = suppforest_redo[!(is.na(suppforest_redo$Type)),]

order = c( 'LPS', 'H1N1')
suppforest_redo$Type = factor(suppforest_redo$Type, levels=order)

order1 = c( 'Male', 'Female')
suppforest_redo$Sex = factor(suppforest_redo$Sex, levels=order1)

dotCOLS = c("#f9b282", "#a6d8f0")
barCOLS = c("#de6b35", "#008fd5")

supptable = ggplot(suppforest_redo, aes(x=Type, y=Estimate, ymin=Lower_CI, ymax=Upper_CI, color = Sex, fill = Sex))+
 geom_linerange(size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS, guide=TRUE)+
scale_color_manual(values=dotCOLS, guide=FALSE)+
scale_x_discrete(name="") +
scale_y_continuous(name="Eosinophil Effect (95% CI)", limits = c(-2, 2)) +
coord_flip() +
theme_classic(base_size=18 ) +
  facet_wrap(~Set, ncol = 1, ) + theme(strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=18), legend.title = element_blank()) + guides(fill = guide_legend(override.aes = list(linetype=0), reverse = TRUE))

ggsave("proanti.pdf")
ggsave("proanti.jpg")

```

Calculate meta-analysis metric

```{r}
library(meta)
# make meta term for H1N1 and LPS separately and also within each sex
H_F = heatall[heatall$Medum == "H1N1" & heatall$Sex == "Female",]
H_M = heatall[heatall$Medum == "H1N1" & heatall$Sex == "Male",]
H_A = heatall[heatall$Medum == "H1N1",]

L_F = heatall[heatall$Medum == "LPS" & heatall$Sex == "Female",]
L_M = heatall[heatall$Medum == "LPS" & heatall$Sex == "Male",]
L_A = heatall[heatall$Medum == "LPS",]

H_F_mod = metagen(Estimate, Std..Error,data=H_F, studlab=paste(Cytokine),comb.fixed = TRUE,comb.random = FALSE, prediction=TRUE,sm="SMD")
H_M_mod = metagen(Estimate, Std..Error,data=H_M, studlab=paste(Cytokine),comb.fixed = TRUE,comb.random = FALSE, prediction=TRUE,sm="SMD")
H_A_mod = metagen(Estimate, Std..Error,data=H_A, studlab=paste(Cytokine),comb.fixed = TRUE,comb.random = FALSE, prediction=TRUE,sm="SMD")

L_F_mod = metagen(Estimate, Std..Error,data=L_F, studlab=paste(Cytokine),comb.fixed = TRUE,comb.random = FALSE, prediction=TRUE,sm="SMD")
L_M_mod = metagen(Estimate, Std..Error,data=L_M, studlab=paste(Cytokine),comb.fixed = TRUE,comb.random = FALSE, prediction=TRUE,sm="SMD")
L_A_mod = metagen(Estimate, Std..Error,data=L_A, studlab=paste(Cytokine),comb.fixed = TRUE,comb.random = FALSE, prediction=TRUE,sm="SMD")

# add meta results to df
met = read.csv("~/Desktop/Tsimane/meta.csv", stringsAsFactors = FALSE)
heatmeta = rbind(heatall, met)

# now remove the overrall ones
heatmeta = heatmeta[!(heatmeta$Sex== "All"),]

order1 = c( 'Male', 'Female')
heatmeta$Sex = factor(heatmeta$Sex, levels=order1)

# order variables
order3 = c('Overall', 'TNF-α', 'GM-CSF', 'IL-13', 'IL-12p70', 'IL-10', 'IL-8', 'IL-7', 'IL-6', 'IL-5', 'IL-4', 'IL-2','IL-1β', 'IFNγ')
heatmeta$Cytokine = factor(heatmeta$Cytokine, levels=order3)

dotCOLS_meta = c("#f9b282", "#a6d8f0")
barCOLS_meta = c("#de6b35", "#008fd5")

ggplot() +
  geom_rect(data=ToothGrowth,xmin = 0, xmax = 1.5, ymin = -7, ymax = 3, fill = '#bbd0f2', alpha = 0.01) +
  geom_rect(data=ToothGrowth,xmin = 2, xmax = 8.5, ymin = -7, ymax = 3, fill = '#e6e8e8', alpha = 0.01) +
  geom_rect(data=ToothGrowth,xmin = 8.5, xmax = 11.5, ymin = -7, ymax = 3, fill = '#fae8c8', alpha = 0.01) + geom_rect(data=ToothGrowth,xmin = 11.5, xmax = 15, ymin = -7, ymax = 3, fill = '#eddff5', alpha = 0.05) +
geom_text(data=data.frame(0),x=0.25,y=-6.5,hjust=0,label='') +
geom_text(data=data.frame(0),x=08,y=-6.5,hjust=0,label='Other') +
geom_text(data=data.frame(0),x=11,y=-6.5,hjust=0,label='Th2') +
geom_text(data=data.frame(0),x=14,y=-6.5,hjust=0,label='Th1') +
geom_linerange(data=heatmeta, aes(x=Cytokine, ymin=Lower_CI, ymax=Upper_CI, color = Sex),size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(data=heatmeta, aes(x=Cytokine, y=Estimate, fill = Sex),size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS_meta, guide=TRUE)+
scale_color_manual(values=dotCOLS_meta, guide=FALSE)+
scale_y_continuous(name="Eosinophil Effect (95% CI)", limits = c(-7, 2), breaks=seq(-6,2,2)) +
theme_classic(base_size=18 ) +
facet_wrap(~Medum, ncol = 2, ) + theme(axis.title.y = element_blank(), strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=20), legend.title = element_blank(),axis.text.x = element_text(vjust =1, hjust=0, size = 16)) + guides(fill = guide_legend(override.aes = list(linetype=0), reverse = TRUE)) + coord_flip() 

```

Circulating stims (non-stimulation)

```{r}

######################### Women ######################### 

summary(glm(data = wom_all, log(ifng) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_1b) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_2) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_4) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_5) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_6) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_7) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_8) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_10) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_12p70) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(il_13) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(gm_csf) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = wom_all, log(tnf_a) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))

######################### Men ######################### 

summary(glm(data = men_all, log(ifng) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_1b) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_2) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_4) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_5) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_6) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_7) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_8) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_10) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_12p70) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(il_13) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(gm_csf) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
summary(glm(data = men_all, log(tnf_a) ~ bmi + stim_age + pregnant + log(eos_st) +  log(c_leucocitos)))
```

Plot serum cytokines 

```{r}
serum = read.csv("~/Desktop/Tsimane/serum_forest.csv", stringsAsFactors = FALSE)

# select only eosinophils
serum = serum[serum$Predictor.Variables == "Eosinophil count (logged)",]

# make meta term for within each sex
S_F = serum[serum$Sex == "Female",]
S_M = serum[serum$Sex == "Male",]

S_F_mod = metagen(Estimate, Std..Error,data=S_F, studlab=paste(Cytokine),comb.fixed = TRUE,comb.random = FALSE, prediction=TRUE,sm="SMD")
S_M_mod = metagen(Estimate, Std..Error,data=S_M, studlab=paste(Cytokine),comb.fixed = TRUE,comb.random = FALSE, prediction=TRUE,sm="SMD")

# add overall meta result to heatall df
serum = read.csv("~/Desktop/Tsimane/serum_forest.csv", stringsAsFactors = FALSE)

# fix cytokine names
serum$Cytokine = ifelse(serum$Cytokine == "IFNg", "IFN\u03B3", serum$Cytokine)
serum$Cytokine = ifelse(serum$Cytokine == "1b", "IL-1\u03B2", serum$Cytokine)
serum$Cytokine = ifelse(serum$Cytokine == "TNF", "TNF-\u03B1", serum$Cytokine)
serum$Cytokine = ifelse(serum$Cytokine == "GM", "GM-CSF", serum$Cytokine)

# select only eosinophils
serum = serum[serum$Predictor.Variables == "Eosinophil count (logged)",]

# define colors and order
dotCOLS_1 = c( "#f9b282","#a6d8f0")
barCOLS_1 = c("#de6b35", "#008fd5" )

order5 = c('Overall', 'TNF-α', 'GM-CSF','IL-12p70', 'IL-10', 'IL-8', 'IL-7', 'IL-6', 'IL-1β','IL-13','IL-5', 'IL-4', 'IL-2', 'IFNγ')
serum$Cytokine = factor(serum$Cytokine, levels=order5)

order4 = c(  'Male', 'Female')
serum$Sex = factor(serum$Sex, levels=order4)

# make categories th1/th2/ambiguous
serum$type = ifelse(serum$Cytokine == "IFNγ" | serum$Cytokine == "IL-2", "Th1", NA)
serum$type = ifelse(serum$Cytokine == "IL-4" | serum$Cytokine == "IL-5" | serum$Cytokine == "IL-13", "Th2", "Ambiv")

ggplot() +
  geom_rect(data=ToothGrowth,xmin = 0, xmax = 1.5, ymin = -7, ymax = 4, fill = '#bbd0f2', alpha = 0.01) +
  geom_rect(data=ToothGrowth,xmin = 2, xmax = 8.5, ymin = -7, ymax = 4, fill = '#e6e8e8', alpha = 0.01) +
  geom_rect(data=ToothGrowth,xmin = 8.5, xmax = 11.5, ymin = -7, ymax = 4, fill = '#fae8c8', alpha = 0.01) + geom_rect(data=ToothGrowth,xmin = 11.5, xmax = 15, ymin = -7, ymax = 4, fill = '#eddff5', alpha = 0.05) +
geom_text(data=data.frame(0),x=0.25,y=-6.5,hjust=0,label='') +
geom_text(data=data.frame(0),x=08,y=-6.5,hjust=0,label='Other') +
geom_text(data=data.frame(0),x=11,y=-6.5,hjust=0,label='Th2') +
geom_text(data=data.frame(0),x=14,y=-6.5,hjust=0,label='Th1') +
geom_linerange(data=serum, aes(x=Cytokine, ymin=Lower_CI, ymax=Upper_CI, color = Sex),size=.8,position=position_dodge(width = 0.5)) +
geom_hline(yintercept=0, lty=2) +
geom_point(data=serum, aes(x=Cytokine, y=Estimate, fill = Sex),size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
scale_fill_manual(values=barCOLS_meta, guide=TRUE)+
scale_color_manual(values=dotCOLS_meta, guide=FALSE)+
scale_y_continuous(name="Eosinophil Effect (95% CI)", limits = c(-7, 2), breaks=seq(-6,2,2)) + 
theme_classic(base_size=18 ) +
facet_wrap(~Medum, ncol = 2, ) + theme(axis.title.y = element_blank(), strip.background=element_blank() , panel.grid.major.x = element_line(size=0.2, color = '#cccccc'), strip.text = element_text(size=20), legend.title = element_blank(),axis.text.x = element_text(vjust =1, hjust=0, size = 16)) + guides(fill = guide_legend(override.aes = list(linetype=0), reverse = TRUE)) + coord_flip() 
```

Parasite prevalence

```{r}

colSums(wom_all["a_lum"], na.rm = TRUE) #14/82 17%
colSums(men_all["a_lum"], na.rm = TRUE) #25/97 25%

colSums(wom_all["t_trich"], na.rm = TRUE) #5/82 6%
colSums(men_all["t_trich"], na.rm = TRUE) #5/97 5%

colSums(wom_all["unc"], na.rm = TRUE) #69/82 84%
colSums(men_all["unc"], na.rm = TRUE) #66/97 68%

summary(glm(a_lum ~ male_stim*stim_age, family = "binomial", data = re_all))
summary(glm(t_trich ~ male_stim*stim_age, family = "binomial", data = re_all))
summary(glm(unc ~ male_stim*stim_age, family = "binomial", data = re_all))
summary(glm(strong ~ male_stim*stim_age, family = "binomial", data = re_all))
```
